Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-21 paper/structureScene.tex
update log compression discussion

The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors. We then train models for these classes which are used to automatically annotate our recordings \cite{}. \ja{What's the reference here?} A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class. This allows the scientist to drastically reduce the amount of information requiring manual processing. However, this approach has two drawbacks. First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted. Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.

@@

The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The distance on this average feature vector is then used to determine $p@k$. For the BOF approach of Aucouturier et al. \cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird}. The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation. To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset. the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$. Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features. Recommended number of Gaussians for MFCCs given in \cite{aucouturier2007bag} is 10.

@@

As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied to the two types of features.

TODO ML comment new figure

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/baselines}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. } \ja{Should we not have $k$ stricly smaller than $10$ here?}}
\label{fig:ASS_-1}
\end{center}
\end{figure}

@@

Figure \ref{fig:ASS_0} shows how logarithmic compression is beneficial irrespective of the similarity model used. The gain is more pronounced when considering alternative similarities. \ja{What's ``alternative'' similarities?}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/log}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_0}
\end{center}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/overall}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_1}
\end{center}
\end{figure}

>>>>

The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors. Computational models are trainedthen  for these classes which are used to automatically annotate recordings \cite{quteprints96243}. A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class. This allows the scientist to drastically reduce the amount of information requiring manual processing. However, this approach has two drawbacks. First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted. Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.

@@

The \emph{RbQ} measures are compared to commonly used early integration approach \emph{early}, which consists in averaging over time the feature set of each scene, resulting in one feature vector per scene. The distance on this average feature vector is then used to determine $p@k$. For the BOF approach of Aucouturier et al. \cite{aucouturier2007bag}, GMMs are estimated for each scene using the expectation-maximization (EM) algorithm \cite{dempster-laird, moon1996expectation}. The similarity between a given pair of scene GMMs is then calculated through Monte Carlo sampling approximation. To ensure convergence of the EM algorithm for the scattering features, we reduce their dimension from $1367$ to $30$ by projecting the features onto the top $30$ principal components of the dataset. the number of Gaussians is optimized for each type of features by grid search in the range $[2, 20]$. Best $p@5$ is reached with 8 and 4  Gaussians, respectively for MFCCs and scattering features. Recommended number of Gaussians for MFCCs given in \cite{aucouturier2007bag} is 10.

@@

As seen in Figure \ref{fig:ASS_-1}, the BOF approach behaves similarly to the early approach when applied both to the MFCCs and the scattering features.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/baselines}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.
\vl{The x-axis label is wrong. precision should be replaced by "rank $k$". The y-axis should be "precision ($p@k$)". It would be good to have the y-axis range from $0\%$ to $100\%$.
The lines should be thicker and follow a color palette of complementary colors (blue vs. orange). The legends should not be abbreviated: scatT -> scattering, mfcc -> MFCC. } \ja{Should we not have $k$ stricly smaller than $10$ here?} \ml{very true, we should stop at k=9 for all figures. We can see it reaches a plateau. mfcc, early pca should be discarded.}}
\label{fig:ASS_-1}
\end{center}
\end{figure}

@@

Figure \ref{fig:ASS_0} shows that logarithmic compression of the scattering features is beneficial. For clarity sake, data is shown  for the \emph{early} approach only, but a equivalent gain is achieved for the relevance-based quantization approaches. %The gain is more pronounced when considering alternative similarities. \ja{What's ``alternative'' similarities?}

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/log}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering with or without logarithmic compression, as a function of the rank $k$.\ml{show only early}}
\label{fig:ASS_0}
\end{center}
\end{figure}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{figures/overall}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.\ml{keep only early approaches, and adapt the text refering to it then.}}
\label{fig:ASS_1}
\end{center}
\end{figure}
